---
title: "Flows of Norwegian scholars"
execute:
  warning: false
  message: false
format:
  html:
    code-fold: true
---


```{r}
#| echo: false
library(sf)
library(spData)
library(countrycode)
library(flowmapblue)
library(flowmapper)
library(tidyverse)
library(shiny)

#getwd()
#max(scopus_country_flows$year)
```

```{r}
#| code-summary: "Show the code"
scopus_country_flows <- read.csv("data_processed/scopus_2024_V1_scholarlymigration_countryflows_enriched.csv")
countries <- spData::world
#glimpse(countries)
```

People are from many places in the world.
```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot() +
  geom_sf(data = countries) +
  geom_sf(data = st_point_on_surface(countries), color = "darkred")
```

The author of this article was born here.
```{r}
#| code-summary: "Show the code"
oslo <- st_sfc(st_point(c(10.7522, 59.9139)), crs = st_crs(countries))
ggplot() +
  geom_sf(data = countries) +
  geom_sf(data = oslo,
          color = "darkred", size = 1)
```

And has lived here (for various amounts of time):
```{r}
#| code-summary: "Show the code"
cities <- data.frame(
  name = c("Grenoble", "Trondheim", "Paris", "Lima", "Rostock"),
  lon = c(5.7245, 10.3951, 2.3522, -77.0428, 12.1308),
  lat = c(45.1885, 63.4305, 48.8566, -12.0464, 54.0887)
)

# Create Oslo point
oslo <- data.frame(
  name = "Oslo",
  lon = 10.7522,
  lat = 59.9139
)

# Convert to sf objects
cities_sf <- st_as_sf(cities, coords = c("lon", "lat"), crs = st_crs(countries))
oslo_sf <- st_as_sf(oslo, coords = c("lon", "lat"), crs = st_crs(countries))

# Plot
ggplot() +
  geom_sf(data = countries) +
  geom_sf(data = cities_sf, color = "darkgreen", size = 1) +
  geom_sf(data = oslo_sf, color = "darkred", size = 1)
```

Migration is a common phenomenon. This (short) article tries to describe the migration of Norwegian scholars (an undoubtedly) interesting group. Please note that the authors of this article (for who knows what reason) is not a part of this group. 

```{r}
#| code-summary: "Show the code"
countries_centroids <- countries |> 
  st_centroid() |> # finds a center of each country'spolygon
  st_coordinates() |>  # extracts numeric coorindates from the POINT geometry
  as.data.frame() |> # converts the matrix to a data.frame 
  setNames(c("lon", "lat")) |> # renames the columns
  cbind(countries) |> # adds back all the columns (except for geometry) form `countries`
  select(iso_a2, lon, lat, name_long) # only selects a few columns that we really need
#glimpse(countries_centroids)
```

```{r}
#| code-summary: "Show the code"
# Aggregate flows across all years (1998-2018)
flows_aggregated <- scopus_country_flows %>%
  group_by(countrynamefrom, countrynameto, iso3codefrom, iso3codeto) %>%
  summarise(
    total_migrations = sum(n_migrations, na.rm = TRUE),
    .groups = 'drop'
  )

# Filter for flows FROM Norway
norway_flows <- flows_aggregated %>%
  filter(countrynamefrom == "Norway")

# Join with centroids to get origin coordinates
# Note: countries_centroids uses iso_a2 (2-letter codes) but flows use iso3code (3-letter)
# We'll need to convert or join differently
norway_flows <- norway_flows %>%
  left_join(
    countries_centroids %>% select(name_long, lon, lat),
    by = c("countrynamefrom" = "name_long")
  ) %>%
  rename(origin_lon = lon, origin_lat = lat)

# Join with centroids to get destination coordinates
norway_flows <- norway_flows %>%
  left_join(
    countries_centroids %>% select(name_long, lon, lat),
    by = c("countrynameto" = "name_long")
  ) %>%
  rename(dest_lon = lon, dest_lat = lat)

# Remove any rows with missing coordinates
norway_flows <- norway_flows %>%
  filter(!is.na(origin_lon) & !is.na(dest_lon))

#glimpse(norway_flows)
```

```{r}
#| code-summary: "Show the code"
base_plot <- ggplot() +
  geom_sf(data = countries, fill = "white", col = "grey80", linewidth = 0.1) +
  theme_void(base_size = 14) +
  labs(
    title = "Migration flows from Norway (1998-2018)", 
    fill = "Total Migrations"
  ) +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5, margin = margin(b = 10)),
    plot.background = element_rect(fill = "white", color = NA),
    legend.position = "right"
  )

od_norway <- norway_flows %>%
  select(
    o = countrynamefrom,      # origin
    d = countrynameto,         # destination  
    value = total_migrations   # flow value
  ) %>%
  mutate(
    o = as.character(o),
    d = as.character(d)
  )

# Prepare nodes data - columns needed: name, x, y
nodes_data <- countries_centroids %>%
  select(
    name = name_long,   # This was the issue - needs to be 'name' not 'id'
    x = lon,
    y = lat
  ) %>%
  mutate(name = as.character(name))

# Create base plot
base_plot <- ggplot() +
  geom_sf(data = countries, fill = NA, col = "grey60", linewidth = 0.05) +
  theme_classic(base_size = 20) +
  labs(title = "Migration Flows from Norway (1998-2018)", 
       subtitle = "", 
       fill = "") +
  theme(
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    panel.background = element_rect(fill = "transparent"),
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  guides(fill = "none")

# Add flowmap
flows_plot <- base_plot %>%
  add_flowmap(
    od = od_norway,
    nodes = nodes_data,
    node_radius_factor = 1,        
    edge_width_factor = 1,      
    arrow_point_angle = 50,          
    node_buffer_factor = 1.5,       
    outline_col = "grey80",
    add_legend = "right",    
    legend_col = "gray20",
    legend_gradient = TRUE,
    k_nodes = 15                  
  )

# Better color gradient
flows_plot <- flows_plot +
  scale_fill_gradient(
    low = "#FABB29", 
    high = "#AB061F",
    labels = scales::comma_format()
  ) 

flows_plot
```

As you can see people, basically just go to the US and the UK and the migration is not that high (BUT IT EXISTS). Let’s add immigration as well. 

```{r}
#| code-summary: "Show the code"
# Aggregate flows TO Norway
flows_to_norway <- scopus_country_flows %>%
  group_by(countrynamefrom, countrynameto, iso3codefrom, iso3codeto) %>%
  summarise(
    total_migrations = sum(n_migrations, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  filter(countrynameto == "Norway")

# Join with centroids for TO Norway flows
flows_to_norway <- flows_to_norway %>%
  left_join(
    countries_centroids %>% select(name_long, lon, lat),
    by = c("countrynamefrom" = "name_long")
  ) %>%
  rename(origin_lon = lon, origin_lat = lat) %>%
  left_join(
    countries_centroids %>% select(name_long, lon, lat),
    by = c("countrynameto" = "name_long")
  ) %>%
  rename(dest_lon = lon, dest_lat = lat) %>%
  filter(!is.na(origin_lon) & !is.na(dest_lon))

# Prepare OD data for TO Norway
od_to_norway <- flows_to_norway %>%
  select(
    o = countrynamefrom,
    d = countrynameto,
    value = total_migrations
  ) %>%
  mutate(
    o = as.character(o),
    d = as.character(d)
  )

# Create base plot
base_plot <- ggplot() +
  geom_sf(data = countries, fill = NA, col = "grey60", linewidth = 0.05) +
  theme_classic(base_size = 20) +
  labs(title = "Migration Flows from and to Norway (1998-2018)",
       fill = "") +
  theme(
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    panel.background = element_rect(fill = "transparent"),
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.subtitle = element_text(hjust = 0.5, size = 14)
  ) +
  guides(fill = "none")

# Add flowmap for FROM Norway (red/orange)
flows_plot <- base_plot %>%
  add_flowmap(
    od = od_norway,
    nodes = nodes_data,
    node_radius_factor = 1,        
    edge_width_factor = 1,      
    arrow_point_angle = 50,          
    node_buffer_factor = 1.5,       
    outline_col = "grey80",
    add_legend = "none",
    k_nodes = 15                  
  )

flows_plot <- flows_plot +
  scale_fill_gradient(/Users/emblavesterdal/Global-flows-and-rates-of-international-migration-of-scholars/ScholarsFlows.qmd
    low = "#FABB29", 
    high = "#AB061F",
    labels = scales::comma_format()
  )

# Add flowmap for TO Norway (blue)
flows_plot <- flows_plot %>%
  add_flowmap(
    od = od_to_norway,
    nodes = nodes_data,
    node_radius_factor = 1,        
    edge_width_factor = 1,      
    arrow_point_angle = 50,          
    node_buffer_factor = 1.5,       
    outline_col = "grey80",
    add_legend = "bottom",
    legend_col = "gray20",
    legend_gradient = TRUE,
    k_nodes = 15                  
  )

flows_plot <- flows_plot +
  scale_fill_gradient(
    low = "#9ECAE1",  # Light blue
    high = "#08519C",  # Dark blue
    labels = scales::comma_format()
  )

flows_plot
```

Yes, you are probably completly blown away right now. Thank you so much.

